### 2019-01-16 manuscript/manuscript.rmd
fixed formatting
moved matrix comparison tomanuscript

```{r preliminaries}
library( tidyverse )
library( magrittr )

@@

```{ r taxon_rectangles }

@@

>>>>

```{r preliminaries}
library( tidyverse )
library( magrittr )
library( viridis )

@@

```{r taxon_rectangles }

@@

## Model matrix comparison

WAG and LG are both fixed exchange matrices. Their differences are largely limitted to a few amino acid changes.

```{r wagvlg}

# Comparison of LG and WAG matrices

library( tidyverse )
library( magrittr )

# Order in iqtree matrix
aa = c("A", "R", "N", "D", "C", "Q", "E", "G", "H", "I", "L", "K", "M", "F", "P", "S", "T", "W", "Y", "V")

# Functional groups adjacent, from https://en.wikipedia.org/wiki/Amino_acid
aa_functional = c( "R", "H", "K", "D", "E", "S", "T", "N", "Q", "C", "G", "P", "A", "V", "I", "L", "M", "F", "Y", "W" )

triange_to_matrix = function( triangle ){
	R_matrix = rep(NA, 400)
	dim( R_matrix ) = c(20,20)

	rownames( R_matrix ) = aa
	colnames( R_matrix ) = aa

	# Though iqtree specifies lower triangle, actually in the correct order for upper triangle as this R command reads it
	R_matrix [ upper.tri( R_matrix ) ] = triangle

	# Now fill in the lower half of the matrix
	R_matrix[ is.na(R_matrix) ] = t( R_matrix )[ is.na(R_matrix) ]

	# Scale
	R_matrix = R_matrix / mean( R_matrix, na.rm=TRUE )

	# Reorder by functional groupings
	R_matrix = R_matrix[, match(aa_functional, colnames(R_matrix) )]
	R_matrix = R_matrix[match(aa_functional, rownames(R_matrix)), ]

	R_matrix
}

plot_matrix = function( R_matrix ){
	D = data.frame( R_matrix )
	D$source = row.names(D)
	Dg = gather( D, 1:20, key=dest, value=exch, -source )
	Dg %<>%
		mutate( source=factor(source, levels=aa_functional)  ) %>%
		mutate( dest=factor(dest, levels=aa_functional)  )

	Dg %>%
		ggplot(aes(x=dest, y=source)) + 
		geom_tile(aes(fill=exch)) +
		scale_fill_gradient2()
}

# Triangle matrices from iqtree source code at https://github.com/Cibiv/IQ-TREE/blob/master/model/modelprotein.cpp

wag_triangle = c(55.15710, 50.98480, 63.53460, 73.89980, 14.73040, 542.94200, 102.70400, 52.81910, 26.52560, 3.02949, 90.85980, 303.55000, 154.36400, 61.67830, 9.88179, 158.28500, 43.91570, 94.71980, 617.41600, 2.13520, 546.94700, 141.67200, 58.46650, 112.55600, 86.55840, 30.66740, 33.00520, 56.77170, 31.69540, 213.71500, 395.62900, 93.06760, 24.89720, 429.41100, 57.00250, 24.94100, 19.33350, 18.69790, 55.42360, 3.94370, 17.01350, 11.39170, 12.73950, 3.04501, 13.81900, 39.79150, 49.76710, 13.15280, 8.48047, 38.42870, 86.94890, 15.42630, 6.13037, 49.94620, 317.09700, 90.62650, 535.14200, 301.20100, 47.98550, 7.40339, 389.49000, 258.44300, 37.35580, 89.04320, 32.38320, 25.75550, 89.34960, 68.31620, 19.82210, 10.37540, 39.04820, 154.52600, 31.51240, 17.41000, 40.41410, 425.74600, 485.40200, 93.42760, 21.04940, 10.27110, 9.61621, 4.67304, 39.80200, 9.99208, 8.11339, 4.99310, 67.93710, 105.94700, 211.51700, 8.88360, 119.06300, 143.85500, 67.94890, 19.50810, 42.39840, 10.94040, 93.33720, 68.23550, 24.35700, 69.61980, 9.99288, 41.58440, 55.68960, 17.13290, 16.14440, 337.07900, 122.41900, 397.42300, 107.17600, 140.76600, 102.88700, 70.49390, 134.18200, 74.01690, 31.94400, 34.47390, 96.71300, 49.39050, 54.59310, 161.32800, 212.11100, 55.44130, 203.00600, 37.48660, 51.29840, 85.79280, 82.27650, 22.58330, 47.33070, 145.81600, 32.66220, 138.69800, 151.61200, 17.19030, 79.53840, 437.80200, 11.31330, 116.39200, 7.19167, 12.97670, 71.70700, 21.57370, 15.65570, 33.69830, 26.25690, 21.24830, 66.53090, 13.75050, 51.57060, 152.96400, 13.94050, 52.37420, 11.08640, 24.07350, 38.15330, 108.60000, 32.57110, 54.38330, 22.77100, 19.63030, 10.36040, 387.34400, 42.01700, 39.86180, 13.32640, 42.84370, 645.42800, 21.60460, 78.69930, 29.11480, 248.53900, 200.60100, 25.18490, 19.62460, 15.23350, 100.21400, 30.12810, 58.87310, 18.72470, 11.83580, 782.13000, 180.03400, 30.54340, 205.84500, 64.98920, 31.48870, 23.27390, 138.82300, 36.53690, 31.47300)
wag_matrix = triange_to_matrix( wag_triangle ) %>% log()

lg_triangle = c(0.425093, 0.276818, 0.751878, 0.395144, 0.123954, 5.076149, 2.489084, 0.534551, 0.528768, 0.062556, 0.969894, 2.807908, 1.695752, 0.523386, 0.084808, 1.038545, 0.363970, 0.541712, 5.243870, 0.003499, 4.128591, 2.066040, 0.390192, 1.437645, 0.844926, 0.569265, 0.267959, 0.348847, 0.358858, 2.426601, 4.509238, 0.927114, 0.640543, 4.813505, 0.423881, 0.311484, 0.149830, 0.126991, 0.191503, 0.010690, 0.320627, 0.072854, 0.044265, 0.008705, 0.108882, 0.395337, 0.301848, 0.068427, 0.015076, 0.594007, 0.582457, 0.069673, 0.044261, 0.366317, 4.145067, 0.536518, 6.326067, 2.145078, 0.282959, 0.013266, 3.234294, 1.807177, 0.296636, 0.697264, 0.159069, 0.137500, 1.124035, 0.484133, 0.371004, 0.025548, 0.893680, 1.672569, 0.173735, 0.139538, 0.442472, 4.273607, 6.312358, 0.656604, 0.253701, 0.052722, 0.089525, 0.017416, 1.105251, 0.035855, 0.018811, 0.089586, 0.682139, 1.112727, 2.592692, 0.023918, 1.798853, 1.177651, 0.332533, 0.161787, 0.394456, 0.075382, 0.624294, 0.419409, 0.196961, 0.508851, 0.078281, 0.249060, 0.390322, 0.099849, 0.094464, 4.727182, 0.858151, 4.008358, 1.240275, 2.784478, 1.223828, 0.611973, 1.739990, 0.990012, 0.064105, 0.182287, 0.748683, 0.346960, 0.361819, 1.338132, 2.139501, 0.578987, 2.000679, 0.425860, 1.143480, 1.080136, 0.604545, 0.129836, 0.584262, 1.033739, 0.302936, 1.136863, 2.020366, 0.165001, 0.571468, 6.472279, 0.180717, 0.593607, 0.045376, 0.029890, 0.670128, 0.236199, 0.077852, 0.268491, 0.597054, 0.111660, 0.619632, 0.049906, 0.696175, 2.457121, 0.095131, 0.248862, 0.140825, 0.218959, 0.314440, 0.612025, 0.135107, 1.165532, 0.257336, 0.120037, 0.054679, 5.306834, 0.232523, 0.299648, 0.131932, 0.481306, 7.803902, 0.089613, 0.400547, 0.245841, 3.151815, 2.547870, 0.170887, 0.083688, 0.037967, 1.959291, 0.210332, 0.245034, 0.076701, 0.119013, 10.649107, 1.702745, 0.185202, 1.898718, 0.654683, 0.296501, 0.098369, 2.188158, 0.189510, 0.249313)
lg_matrix = triange_to_matrix( lg_triangle ) %>% log()

plot_matrix(wag_matrix)
plot_matrix(lg_matrix)

lg_wag_diff = lg_matrix - wag_matrix

plot_matrix(lg_wag_diff)

lg_wag_diff_thresh =lg_wag_diff
lg_wag_diff_thresh[abs(lg_wag_diff_thresh)<1] = 0

plot_matrix( lg_wag_diff_thresh )
